On Adding Pattern Matching to Haskell-based Deeply Embedded Domain Specific Languages

### 2020-10-10 paper/paper.tex
Remove a couple of widows
vertical spacing around figure 1
formatting

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\linenumbers
\begin{abstract}
% Say what problem is
  Capturing control flow is the Achilles heel of Haskell-based
  deeply embedded domain specific languages.
  Rather than use
  the builtin control flow mechanisms, artificial control flow combinators
  are used instead.
% Say why it is an interesting problem
  However, capturing traditional control flow in an deeply embedded domain specific language
  would support the writing of programs in a natural style by allowing the programmer to use the
  constructs that are already builtin to the base language, such as pattern
  matching and recursion.
% Say what our solution achieves
  In this paper, we expand the capabilities of
  Haskell-based deep embeddings with a compiler extension
  for reifying conditionals and pattern matching.
% Here is why the world will be a better place
  With this new support, the Haskell that we capture for expressing
  deeply embedded domain specific languages can be cleaner, Haskell-idiomatic,
  and more declarative in nature.

@@

Given this code, \ttt{ERepTy ComplexPair \typeeq (Complex Double, Complex Double)}. Here
is an example that demonstrates why this is more useful than if it \textit{fully} deconstructed
each type into \ttt{Either}, \ttt{(,)} and \ttt{()}:

@@

% \newpage
\begin{lstlisting}
data E t where
  ...
  SumMatchExp :: (ERep a, ERep b, ERepTy b ~ b) =>
    E (ProdMatch a r) -> E (SumMatch b r)
      -> E (SumMatch (Either a b) r)

  OneSumMatch ::
    (ERep a, ERep b, ERepTy a ~ a) =>
      E (ProdMatch a b) -> E (SumMatch a b)

  EmptyMatch :: (ERep b) => E (SumMatch Void b)
  ...
\end{lstlisting}

@@

\begin{lstlisting}
data E t where
  ...
  Lam :: (ERep a, Typeable a) =>
    Name a -> E b -> E (a -> b)
  Var :: (Typeable a) => Name a -> E a
\end{lstlisting}

@@

\begin{wrapfigure}{r}{0.5\linewidth}
  % \centering
  \resizebox{0.5\columnwidth}{7cm}{\includegraphics{Diagrams/Flowchart.pdf}}
  % \includegraphics{Diagrams/Flowchart.pdf}
   \caption{Control Flow of the Core Plugin}
   \label{fig:CorePlugin}
\end{wrapfigure}%

@@

Note that:
\begin{itemize}
  \item The tail recursion transformation given by $D$, $T'_{\ttt{f}}$ and $T_{\ttt{f}}$ is completely independent of the \ttt{E} type and the transformations associated to the \ttt{E} type. As a result, the tail recursion transformation can be used on its own.

@@

@@

% \captionsetup{justification=centering}
\caption{Rewrite rules}
\label{fig:Rewrites}
% \end{wrapfigure}
% \end{wrapfigure}
\end{figure}

>>>>

%%
%% The abstract is a short summary of the work to be presented in the
%% article.

% \linenumbers

\begin{abstract}
% Say what problem is
  Capturing control flow is the Achilles heel of Haskell-based
  deeply embedded domain specific languages.
  Rather than use
  the builtin control flow mechanisms, artificial control flow combinators
  are used instead.
% Say why it is an interesting problem
  However, capturing traditional control flow in an deeply embedded domain specific language
  would support the writing of programs in a natural style by allowing the programmer to use the
  constructs that are already builtin to the base language, such as pattern
  matching and recursion.
% Say what our solution achieves
  In this paper, we expand the capabilities of
  Haskell-based deep embeddings with a compiler extension
  for reifying conditionals and pattern matching.
% Here is why the world will be a better place
  With this new support, the Haskell that we capture for expressing
  deeply embedded domain specific languages can be cleaner, Haskell-idiomatic,
  and more declarative in nature.

@@

Given this code, the following type equality holds:
\\

\ttt{ERepTy ComplexPair \typeeq (Complex Double, Complex Double)}
\\

Here is an example that demonstrates why this is more useful than if it \textit{fully} deconstructed
each type into \ttt{Either}, \ttt{(,)} and \ttt{()}:

@@

% \newpage
\begin{lstlisting}
data E t where
  ...
  SumMatchExp :: (ERep a, ERep b, ERepTy b ~ b) =>
    E (ProdMatch a r) -> E (SumMatch b r)
      -> E (SumMatch (Either a b) r)
  OneSumMatch ::
    (ERep a, ERep b, ERepTy a ~ a) =>
      E (ProdMatch a b) -> E (SumMatch a b)
  EmptyMatch :: (ERep b) => E (SumMatch Void b)
\end{lstlisting}

@@

\begin{lstlisting}
data E t where
  ...
  Lam :: (ERep a, Typeable a) =>
    Name a -> E b -> E (a -> b)

  Var :: (Typeable a) => Name a -> E a
\end{lstlisting}

@@

@@

% \begin{wrapfigure}{r}{0.5\linewidth}
\begin{figure}
  \centering
  % \vspace{-1cm}
  \resizebox{0.8\columnwidth}{8.2cm}{\includegraphics{Diagrams/Flowchart.pdf}}
  % \includegraphics{Diagrams/Flowchart.pdf}
   \caption{Control Flow of the Core Plugin}
   \label{fig:CorePlugin}
  % \vspace{-0.1cm}
\end{figure}%

\pagebreak
Note that:
\begin{itemize}
  \item The tail recursion transformation given by $D$, $T'_{\ttt{f}}$ and $T_{\ttt{f}}$ is completely independent of the \ttt{E} type and the transformations associated to the \ttt{E} type. As a result, the tail recursion transformation can be used on its own.

@@

  % \vspace{-0.9cm}

@@

% \captionsetup{justification=centering}
\caption{Rewrite rules}
\label{fig:Rewrites}
% \end{wrapfigure}
% \end{wrapfigure}

% \vspace{-1cm}

\end{figure}
